---
title: "Reproducible Data Project 1"
author: "Dorie"
date: "June 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Data Project 1
## Loading and Proccessing Data

```{r}
data <- read.csv("activity.csv")
```
# What is the mean total number of steps taken per day?
# Calculate the total number of steps taken by day and make a histogram of the total number of steps:

```{r}
stepsbyday <- aggregate(steps ~ date, data, sum)
hist(stepsbyday$steps, main = paste("Total Steps by Day"), col= "red", xlab= "Number of Steps")
```
#Calculate the mean and median of the total number of steps taken per day

```{r}
mean(stepsbyday$steps)
median(stepsbyday$steps)
```

# What is the average daily activity pattern?

```{r}
averagenumberofsteps <- aggregate(steps ~ interval, data, mean)
```
#Make a time series plot

```{r}
library("ggplot2")
ggplot(data=averagenumberofsteps, aes(x=interval, y=steps)) +
  geom_line() +
  xlab("5 Minute Intervals") +
  ylab("Average Number of Steps Taken")
```

#Which 5 minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
averagenumberofsteps[which.max(averagenumberofsteps$steps),]
```
# Imputing Missing Values
##Calculate and report the total number of missing values in the dataset:

```{r}
sum(is.na(data$steps))
```
##Fill in all of the missing values in the dataset and create a new dataset that is equal to the original dataset but with missing values filled in:

```{r}
alldata <- data
nas <- is.na(alldata$steps)
avg_interval <- tapply(alldata$steps, alldata$interval, mean, na.rm=TRUE, simplify=TRUE)
alldata$steps[nas] <- avg_interval[as.character(alldata$interval[nas])]

sum(is.na(alldata$steps))

incomplete <- sum(!complete.cases(data))
imputed_data <- transform(data, steps = ifelse(is.na(data$steps),averagenumberofsteps$steps[match(data$interval, averagenumberofsteps$interval)], data$steps))

allstepsbyday <- aggregate(steps ~ date,imputed_data, sum)


ggplot(allstepsbyday, aes(x = steps)) +
  geom_histogram(fill = "red", binwidth = 1000) +
  labs(title = "Histogram of All Steps per Day", x = "Steps per Day", y = "Frequency")
```
##Calculate the new mean and median for imputed data:

```{r}
mean(allstepsbyday$steps)
median(allstepsbyday$steps)
```



